Correlation among smart grid actors

### 2015-07-28 Article.tex
update

Achieving a successful energetic transition through a smarter and greener electricity grid is a major goal for the 21st century. It is assumed that such smart grids will be characterized by bidirectional electricity flows coupled with a large penetration of small renewable generators and a proper efficient information system. All these bricks might enable end users to take part in the grid stability by injecting power, or by shaping their consumption against financial compensation. In this paper, we study the formation of coalitions of these futuristic agents that both produce and consume. These coalitions, which can be seen as aggregators of the electricity market, aim at selling the aggregated surplus of production of the agents they control. We rely on real weather data sampled accross stations of a given territory in order to simulate realistic production and consumption patterns for each agent. This approach enables us to capture geographical correlations among the agents while preserving the diversity due to different behaviors. As aggregators are bound to the grid operator by a contract, they seek to maximize their offer while minimizing their risk. We propose here an algorithm that takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability. We show then that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.
\end{abstract}

@@

@@

\begin{figure}
\includegraphics[scale=.45]{Fig2.pdf}
\caption{Process diagram}
\label{fig:process}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=.6]{gaussian_P_star.pdf}
\caption{$ P_{S}^{CRCT \star} $ depending on reliability parameter $ \phi $ for gaussian distributions (see equation \ref{eq:P_star}). Blue curve with triangles stands for a coalition S with an expected production of 5 units and a standard deviation of 0.5. Under a grid policy of $ \phi = 0.1 $, it is able to announce a contract value of $ P_{S}^{CRCT \star} = 4.36 $. The same coalition in term of expected production ($\mu = 5 $), but with a higher variance ($ \sigma = 1.5 $, cyan curve with circles) can only afford a smaller contract value of $ P_{S}^{CRCT \star} = 3.07 $. The red curve with diamonds stands for a coalition with a higher expected production $(\mu = 7)$, but with a very high unpredictability ($ \sigma = 5 $). For low values of $ \phi $, this coalition is thus heavily penalized and can only afford a contract of $ 0.59 $ units. Under grid policy $ (\phi = 0.1, P^{MIN} = 2) $, this last coalition is thus not allowed to enter the market (red dot below the horrizontal dashed line). }
\label{fig:Gaussian}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=.6]{mean_field.pdf}
\caption{Gaussian mean approximation. Subplot \textbf{a} shows how the parameter $ \alpha $ of the utility function should be chosen in function of the mean desired size of the coalitions (see equation \ref{eq:alpha_star}). Subplot \textbf{b} dislays the corresponding utility functions for different values of $ \alpha $. Blue curve with diamonds favors very small coalitions of 2 agents while the green one with triangles favors 5 agents coalitions. Finally, the red curve with squares has an optimal size of 15 agents. }
\label{fig:mean_approx}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=.6]{cliques_histo.pdf}
\caption{Histograms of utility values for coalitions of size 3. Red bars stands for cliques in the decorrelation graph, and blue bars represents all the other possible triplets. As visible, cliques tend to exhibit higher utilities than randomly selected coalitions.}
\label{fig:histo_cliques}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=.6]{real_utility_2.pdf}
\caption{Utility of random coalitions depending on their size. Blue dots shows real mean utility values and the thick red curve its smoothed version by applying a Savitzky-Golay filter. On this plot the $ \alpha $ parameter of the utility function was selected according to equation \ref{eq:alpha_star} in order to favor 40 agents coalitions.}
\label{fig:real_utility2}
\end{figure}

\begin{figure}
\includegraphics[scale=.6]{search.pdf}
\caption{Evolution of the global utility $ \mathcal{U}(CS) $ (blue diamond curve, left axis) and the number of agents involved in the coalitions (red circle curve, right axis) during the greedy optimization of algorithm \ref{alg:algo1} }
\label{fig:search}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=.6]{./figures/coalitions/coalitions6.pdf}
\caption{Coalitions formed in the (contract value, volatility) space. The color map indicates qualities of portions of the plane. The closer to red the better (high contract values with small volatility). On the opposite, blue areas shows poor quality (small contract values with high volatility). Blue dots stand for the decorrelated coalitions that we formed while green squares show correlated coalitions. The smaller yellow markers stands for the gravity centers of the coalition structures. The black dotted line shows how contract values and volatility evolve  when the size of the coalitions increase (a subset of the points are labeled by the size of the coalition they represent). Each point is the average over $ 10^{5} $ unconstrained draws of a random coalition. As we can also draw random coalition structures, we show the distribution of their gravity centers by the red ellipses (center is the mean, and each ellipse add 1 standard deviation). For comparison with our algorithm where the utility function favors balanced coalitions, we also constrained loosely the sizes of the coalitions in the random process. }
\label{fig:coalitions}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=.6]{./figures/resilience_both.pdf}
\caption{Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot) }
\label{fig:resilience}
\end{figure}

>>>>

Achieving a successful energetic transition through a smarter and greener electricity grid is a major goal for the 21st century. It is assumed that such smart grids will be characterized by bidirectional electricity flows coupled with a large penetration of small renewable generators and a proper efficient information system. All these bricks might enable end users to take part in the grid stability by injecting power, or by shaping their consumption against financial compensation. In this paper, we propose an formation of coalitions algorithm of agents that both produce and consume. Our algorithm is designed to be used by an aggregator in order to form coalitions that aim at selling the aggregated surplus of production of the prosumer it controls. We rely on real weather data sampled across stations of a given territory in order to simulate realistic production and consumption patterns for each agent. This approach enables us to capture geographical correlations among the agents while preserving the diversity due to different behaviors. As aggregators are bound to the grid operator by a contract, they seek to maximize their offer while minimizing their risk. We propose here a graph based algorithm that takes the underlying correlation structure of the agents into account and outputs coalitions with both high productivity and low variability. We show then that the resulting diversified coalitions are able to generate higher benefits on a constrained energy market, and are more resilient to random failures of the agents.

%which can be seen as of the electricity market, .
\end{abstract}

@@

%%%%%%s
Rather than maximizing the profit of each agents in the grid (prosumer and grid operator) our algorithm is trying to find the right trade-off between the productivity and volatility of the power injected into the grid by coalitions of prosumers. However, the pricing strategies for both the prosumer and the grid is beyond the scope of this paper.  
%%%%%%

@@

\begin{figure}
\includegraphics[scale=.45]{Fig2}
\caption{Process diagram}
\label{fig:process}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=.6]{gaussian_P_star}
\caption{$ P_{S}^{CRCT \star} $ depending on reliability parameter $ \phi $ for gaussian distributions (see equation \ref{eq:P_star}). Blue curve with triangles stands for a coalition S with an expected production of 5 units and a standard deviation of 0.5. Under a grid policy of $ \phi = 0.1 $, it is able to announce a contract value of $ P_{S}^{CRCT \star} = 4.36 $. The same coalition in term of expected production ($\mu = 5 $), but with a higher variance ($ \sigma = 1.5 $, cyan curve with circles) can only afford a smaller contract value of $ P_{S}^{CRCT \star} = 3.07 $. The red curve with diamonds stands for a coalition with a higher expected production $(\mu = 7)$, but with a very high unpredictability ($ \sigma = 5 $). For low values of $ \phi $, this coalition is thus heavily penalized and can only afford a contract of $ 0.59 $ units. Under grid policy $ (\phi = 0.1, P^{MIN} = 2) $, this last coalition is thus not allowed to enter the market (red dot below the horrizontal dashed line). }
\label{fig:Gaussian}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=.6]{mean_field}
\caption{Gaussian mean approximation. Subplot \textbf{a} shows how the parameter $ \alpha $ of the utility function should be chosen in function of the mean desired size of the coalitions (see equation \ref{eq:alpha_star}). Subplot \textbf{b} dislays the corresponding utility functions for different values of $ \alpha $. Blue curve with diamonds favors very small coalitions of 2 agents while the green one with triangles favors 5 agents coalitions. Finally, the red curve with squares has an optimal size of 15 agents. }
\label{fig:mean_approx}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=.6]{cliques_histo}
\caption{Histograms of utility values for coalitions of size 3. Red bars stands for cliques in the decorrelation graph, and blue bars represents all the other possible triplets. As visible, cliques tend to exhibit higher utilities than randomly selected coalitions.}
\label{fig:histo_cliques}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=.6]{real_utility_2}
\caption{Utility of random coalitions depending on their size. Blue dots shows real mean utility values and the thick red curve its smoothed version by applying a Savitzky-Golay filter. On this plot the $ \alpha $ parameter of the utility function was selected according to equation \ref{eq:alpha_star} in order to favor 40 agents coalitions.}
\label{fig:real_utility2}
\end{figure}

\begin{figure}
\includegraphics[scale=.6]{search}
\caption{Evolution of the global utility $ \mathcal{U}(CS) $ (blue diamond curve, left axis) and the number of agents involved in the coalitions (red circle curve, right axis) during the greedy optimization of algorithm \ref{alg:algo1} }
\label{fig:search}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=.6]{./figures/coalitions/coalitions6}
\caption{Coalitions formed in the (contract value, volatility) space. The color map indicates qualities of portions of the plane. The closer to red the better (high contract values with small volatility). On the opposite, blue areas shows poor quality (small contract values with high volatility). Blue dots stand for the decorrelated coalitions that we formed while green squares show correlated coalitions. The smaller yellow markers stands for the gravity centers of the coalition structures. The black dotted line shows how contract values and volatility evolve  when the size of the coalitions increase (a subset of the points are labeled by the size of the coalition they represent). Each point is the average over $ 10^{5} $ unconstrained draws of a random coalition. As we can also draw random coalition structures, we show the distribution of their gravity centers by the red ellipses (center is the mean, and each ellipse add 1 standard deviation). For comparison with our algorithm where the utility function favors balanced coalitions, we also constrained loosely the sizes of the coalitions in the random process. }
\label{fig:coalitions}
\end{figure}

@@

\begin{figure}
\includegraphics[scale=.6]{./figures/resilience_both}
\caption{Resilience of the coalition structures when nodes fail randomly (see equation \ref{eq:resilience}) for $ P^{MIN} = 10MW $ (top subplot) and $ P^{MIN} = 80MW $ (bottom subplot) }
\label{fig:resilience}
\end{figure}
